---
title: "MT Cars Visualization"
author: "Mike Calabro"
execute:
  echo: True
  message: False
  warning: False
---

## Purpose

The purpose of this project is to perform Exploratory Data Analysis on the mtcars data set. I am less curious about searching for / communicating insights, and more-so focused on making good-looking graphs and tables :)

## The Data

From the RDocumentation: "The data was extracted from the 1974 *Motor Trend* US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973\--74 models)."

::: {style="font-size:14px"}
```{r}
library(tidyverse)
library(reactable)

data <- mtcars %>% 
  as_tibble(rownames="model")

data %>%
  reactable(
    defaultPageSize = 5, 
    striped = TRUE,
    defaultColDef = colDef(minWidth = 55),
    columns = list(model = colDef(minWidth = 165))
  )
```
:::

## Do Heavier Cars Get Less MPG?

```{r}
data %>%
  ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  labs(x = "Weight (Tons)", y = "MPG", title = "Weight vs. Fuel Eficiency") +
  geom_smooth(method = "loess", se = FALSE) +
  theme_minimal() +
  theme(text = element_text(size = 18),
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5, face = "bold"))
```

### What if we Incorporate Engine Cylinders?

```{r}
mtcars %>%
  ggplot(aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point() +
  geom_smooth(aes(color = factor(cyl)),method = "lm", se = FALSE, alpha = 0.5) +
  labs(color = "Cyl", title = "Car Fuel Economy", x = "Weight", y = "MPG") +
  theme_minimal() +
  theme(text = element_text(size = 18),
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.position = c(0.8,0.76),
        legend.box.background = element_rect(fill = "white", color = "black"),
        axis.text.x = element_text(vjust = 3),
        axis.text.y = element_text(hjust = 1.5))
```

## Comparing Brands

If we wanted to compare brands, a first good step would be to create separate columns for the make and model of a car.

```{r}
library(stringr)

make_model <- data %>%
  mutate(make = ifelse(
                  is.na(str_locate(model, " ")[,"start"]),
                  model,
                  str_sub(model, end = str_locate(model, " ")[,"start"])
                ))

make_model %>%
  count(make) %>%
  ggplot(aes(x = reorder(make, -n), y = n)) +
  geom_col() +
  labs(x = "Brand", y = "Count", title = "Car Frequency in MTCars Data Set") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1),
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", hjust = 0.5))
```

## Are Horse Power and Carborators Independent?

```{r}
make_model %>%
  ggplot(aes(x = factor(carb), y = hp, color = factor(carb))) +
  geom_violin(draw_quantiles = 0.5) +
  geom_jitter(width = 0.05, alpha = 0.8) +
  labs(x = "Carborators", y = "Horse Power", title = "Carbs to HP") +
  theme_minimal() +
  theme(text = element_text(size = 18),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", hjust = 0.5))
```
