---
title: "Mike's Projects"
execute: 
  echo: FALSE
  warning: FALSE
  message: FALSE
---

```{r}
library(tidyverse)
library(janitor)
library(scales)
```

<br>

::: columns
::: {.column width="40%"}
```{r}
#| fig-height: 6

pbp_22 <- readRDS("~/Documents/RStudio/data_science_portfolio/data/nfl_22.rds")

pbp_22 %>%
  filter(play_type == "run") %>%
  ggplot(aes(x = yards_gained)) +
  geom_histogram(fill = "white", color = "blue") +
  labs(x = "Rushing Yards", y = "Count", title = "2022 NFL Rushing Yards") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(limits = c(-10, 50)) +
  theme_minimal() +
  theme(text = element_text(size = 20),
        plot.background = element_rect(color = "black"),
        plot.margin = margin(10,10,10,10),
        axis.text.x = element_text(vjust = 3))

```
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
[**nflfastR Introduction**]{style="font-size: 26px"}

*March 5, 2023*

This project intorduces NFL play-by-play data acquired using the nflfastR package.

[Check out the project here](nflfastR_intro.html)!
:::
:::

<hr>

::: columns
::: {.column width="55%"}
[**Motor Trend Cars Visualization**]{style="font-size: 26px"}

*March 3, 2023*

This project is a short investigation of the MTCars data set. The MTCars data set describes various features of 32 cars, such as miles per gallon, horsepower, and weight.

[Check out the project here](mt_cars_viz.html)!
:::

::: {.column width="5%"}
:::

::: {.column width="40%"}
```{r}
#| fig-height: 6

mtcars %>%
  ggplot(aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point() +
  geom_smooth(aes(color = factor(cyl)),method = "lm", se = FALSE, alpha = 0.5) +
  labs(color = "Cyl", title = "Car Fuel Economy", x = "Weight", y = "MPG") +
  scale_y_continuous(limits = c(0,35)) +
  theme_minimal() +
  theme(text = element_text(size = 20),
        plot.background = element_rect(color = "black"),
        plot.margin = margin(10,10,10,10),
        legend.position = c(0.8,0.76),
        legend.box.background = element_rect(fill = "white", color = "black"),
        axis.text.x = element_text(vjust = 3),
        axis.text.y = element_text(hjust = 1.5))
```
:::
:::

<hr>

::: columns
::: {.column width="40%"}
```{r}
#| fig-height: 6

iris %>%
  clean_names() %>%
  ggplot(aes(x = species, y = petal_width, color = species)) +
  geom_boxplot() +
  labs(title = "Iris Petal Comparisson", x = "Species", y = "Petal Width") +
  scale_y_continuous(limits = c(0,2.7), n.breaks = 6) +
  theme_minimal() +
  theme(text = element_text(size = 20),
        plot.background = element_rect(color = "black"),
        plot.margin = margin(10,10,10,10),
        legend.position = "none",
        legend.box.background = element_rect(fill = "white", color = "black"),
        axis.text.x = element_text(vjust = 2),
        axis.text.y = element_text(hjust = 1.5))
```
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
[**Iris Visualization**]{style="font-size: 26px"}

*February 28, 2023*

This project is a short investigation of the Iris data set. The Iris data set describes 150 samples of 3 species of iris flower in terms of their sepal and petals.

[Check out the project here](iris_viz.html)!
:::
:::
